### 2022-11-25 paper1.tex
third paper
paper 3 supplement
paper 2 start
chapter specific referencing

This paper is structured as follows: Section \ref{sec:methods} presents the
derivation of the sceptical Bayes factor. Its asymptotic and finite sample
properties are then compared with other measures of replication success in
Section \ref{sec:comparison}. An extension to non-normal models is presented in
Section \ref{sec:tdist}. Section \ref{sec:applications} illustrates how the
method works in practice using case studies from the \emph{Social Sciences
  Replication Project} \citep{Camerer2018}. Section \ref{sec:discussion}
provides concluding remarks about strengths, limitations and extensions of the
method.

@@

We adopt a meta-analytic framework and consider the effect estimates as the
data, rather than their underlying samples, and assume that
$\that_k \given \theta \sim \Nor(\theta, \sigma^2_k)$ for $k \in \{o, r\},$ \ie
normality of the effect estimates around $\theta$, with known variances equal to
their squared standard errors. For studies with reasonable sample size, this
framework usually provides a good approximation for a wide range of (suitably
transformed) effect size types \citep[Chapter 2.4]{Spiegelhalter2004}. For
example, means and mean differences (no transformation), odds ratios, hazard
ratios, risk ratios (logarithmic transformation), or correlation coefficients
(``Fisher-$z$'' transformation). We refer to the literature of meta-analysis for
details about transformations of effect sizes \citep[\eg][Chapter
11.6]{Cooper2019}. The normal model in combination with conjugate priors enables
derivation of closed-form expressions in many cases, which allows us to easily
study limiting behaviour and facilitates interpretability. In Section
\ref{sec:tdist}, we will present relaxations of the normality assumption, which
can lead to more accurate inferences when studies have small sample sizes and/or
show extreme results.

\section{Reverse-Bayes assessment of replication success with Bayes factors}
\label{sec:methods}
% ``judgements are made concerning the values that the final
% probabilities would have after some imaginary results are obtained; then Bayes'
% theorem is used backwards to estimate what the consistent initial probabilities
% must be assumed to be; and lastly Bayes's theorem is used (forwards) to obtain
% estimates of the final probabilities in the light of the actual observations.
% This device takes \emph{seriously} the view that a theory of probability or of
% rationality is no more than a theory of consistency'' (p. 126).
The idea of reversing Bayes' theorem was first proposed by \citet{Good1950}. He
acknowledged that in many situations there is no obvious choice for the prior
distributions involved in Bayesian analyses. On the other hand, we are often
more certain which posterior inferences would convince us regarding the
credibility of a hypothesis. For this reason, Good inverted Bayes' theorem and
derived priors, which combined with the observed data, would lead to posterior
inferences that were specified beforehand (\eg the data favour one hypothesis
over another). His reverse-Bayes inference then centred around the question
whether the resulting prior is plausible, and if so, this would legitimise the
posterior inference. See Figure~\ref{fig:RBschmema} for a graphical illustration
of this process.

\begin{figure}[!htb]
  \centering
%% first attempt
\begin{tikzpicture}[node distance = 5em, thick]
  % nodes %
  \node[font = \large] (data) {$f(\mathrm{data} \,|\, \theta)$};
  \node[font = \large] (prior) [left=7em of data] {$f(\theta)$};
  \node[font = \large] (post) [right=7em of data] {$f(\theta \,|\, \mathrm{data})$};
  \node[text width=10em,text centered] (RBinference) [below of=prior] {Reverse-Bayes inference};
  \node[text width=10em, text centered] (FBinference) [below of=post] {Forward-Bayes inference};

  % edges %
  \draw [->, dotted] (prior) to [out=45, in=135, looseness=0.3] node[above]{Bayesian updating} (post);
  %% other names could be:
  %% - reverse Bayesian updating vs forward Bayesian updating
  %% - Bayesian reverting
  \draw [->, dashed] (post) to [out=225, in=315, looseness=0.3] node[below]{Bayesian downdating} (prior);
  \draw [->, dashed] (prior) to [out=270, in=90, looseness=0] (RBinference);
  \draw [->, dotted] (post) to [out=270, in=90, looseness=0] (FBinference);
\end{tikzpicture}
\caption{Schematic illustration of reverse-Bayes and forward-Bayes inference
  for an unknown parameter $\theta$.}
\label{fig:RBschmema}
\end{figure}

@@

\subsection{Data from the original study}
For the effect estimate $\that_o \given \theta \sim \Nor(\theta, \sigma^2_o)$
from the original study consider a hypothesis test comparing the null hypothesis
$H_0\colon$ $\theta = 0$ to the alternative $\HS \colon$ $\theta \neq 0$.
Specification of a prior distribution for $\theta$ under $\HS$ is now required
for Bayesian hypothesis testing. A typical choice \citep{Jeffreys1961} is a
local alternative, a unimodal symmetric prior distribution centred around the
null value. We consider the sceptical prior
$\theta \given \HS \sim \Nor(0, \sigma^2_s = g \cdot \sigma^2_o)$ with relative
sceptical prior variance $g$ for this purpose (relative to the variance from the
original estimate $g = \sigma^2_s/\sigma^2_o$), resembling the $g$-prior known
from the regression literature \citep{Zellner1986, Liang2008}. The explicit form
of the Bayes factor is then given by
\begin{align}
  \label{eq:BFo}
  \BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)
  = \sqrt{1 + g} \cdot
  \exp\left\{-\frac{1}{2} \cdot \frac{g}{1 + g} \cdot z^2_o\right\}.
\end{align}

The Bayes factor from equation ~\eqref{eq:BFo} is shown in Figure ~\ref{fig:bfo}
as a function of $g$ and for different original $z$-values $z_o$. For fixed
$z_o$, it is well known that this Bayes factor is bounded from below by
\begin{align}
  \label{eq:minBFo}
  \minBF_{o} =
  \begin{cases}
    |z_o| \cdot  \exp(-z_o^2/2) \cdot \sqrt{e} & \text{for} ~ |z_o| > 1 \\
    1 & \text{for} ~ |z_o| \leq 1
  \end{cases}
\end{align}
which is reached at $\gminBFo = \max\{0, z^2_o - 1\}$ \citep{Edwards1963}.
Further increasing the relative sceptical prior variance increases
\eqref{eq:BFo} indefinitely because of the Jeffreys-Lindley paradox, \ie
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g) \to \infty$ for $g \to \infty$
\citep[][Section 6.1.4]{Bernardo2000}. Hence, for a relative sceptical prior
variance $g \in [0, \gminBFo]$, the resulting Bayes factor will be
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g) \in [\minBF_o, 1]$.

\begin{figure}[!htb]
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{images/paper1/plot-BFo-1.pdf}
\end{knitrout}
\caption{Bayes factor $\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ as a function of
relative sceptical prior variance $g$ for different values of
$|z_o| = |\that_o|/\sigma_o$. Minimum Bayes factors $\minBF_{o}$ are indicated
by dots. Dashed vertical lines indicate sufficiently sceptical relative prior
variance $g_{\scriptscriptstyle \gamma}$ at level $\gamma = 1/10$, if they
exist.}
\label{fig:bfo}
\end{figure}

@@

From Figure \ref{fig:bfo} we can see that the more compelling the original data
(\ie the larger $|z_o|$), the smaller the sufficiently sceptical relative prior
variance $\gy$ needs to be in order to make the result no longer convincing at
level $\gamma$. In the most extreme case, when $|z_{o}| \to \infty$ and $\gamma$
remains fixed, the sufficiently sceptical prior variance will converge to
zero (Appendix \ref{appendix:lambertW}). On the other hand, if $|z_o|$ is not
sufficiently large, $\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ will either be
always increasing in $g$ (if $|z_o| \leq 1$) or it will reach a minimum above
the chosen level $\gamma$. In both cases the sufficiently sceptical relative
prior variance $\gy$ is not defined since there is no need to challenge an
already unconvincing result.

@@

\subsection{Data from the replication study}
\label{sec:bfr}
In order to assess whether the original finding can be replicated in an
independent study, a replication study is conducted, leading to a new effect
estimate $\hat{\theta}_r$. In light of the new data, the sceptic is now
challenged by an advocate of the original finding. This is formalised with
another Bayes factor, which compares the plausibility of the replication effect
estimate $\hat{\theta}_r \given \theta \sim \Nor(\theta, \sigma^2_r)$ under the
sceptical prior $\HS \colon$ $\theta \sim \Nor(0, g \cdot \sigma^2_o)$ relative
to the advocacy prior $\HA \colon$ $\theta \sim \Nor(\that_o, \sigma^2_o)$. The
the view of an advocate is represented by $\HA$ since this is the posterior of
$\theta$ given the original estimate and a uniform prior (also the reference
prior for this model). The Bayes factor is given by
\begin{align}
  \label{eq:BFr}
  \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g)
%%   &= \sqrt{\frac{1/c + 1}{1/c + g}} \cdot
%%   \exp\left\{-\frac{1}{2}\left(\frac{z_r^2}{1/c + g} -
%%   \frac{\left(z_r - z_o \sqrt{c}\right)^2}{1 + c} \right)\right\}
  &= \sqrt{\frac{1/c + 1}{1/c + g}} \cdot
  \exp\left\{-\frac{z_o^2}{2}\left(\frac{d^2}{1/c + g} -
  \frac{\left(d - 1\right)^2}{1/c + 1} \right)\right\}
\end{align}
so it depends on the original $z$-statistic $z_o$, the relative sceptical prior
variance $g$, the relative effect estimate $d = \that_r/\that_o$, and the
relative variance $c = \sigma^2_o/\sigma^2_r$.

Our goal is now to define a condition for \emph{replication success} in terms of
\eqref{eq:BFr}. It is natural to consider a replication successful if the
replication data favour the advocate over the sceptic to a higher degree than
the sceptic's initial objection to the original study. More formally, we say
that if the Bayes factor from \eqref{eq:BFr} evaluated at the sufficiently
sceptical relative prior variance $\gy$ is not larger than the corresponding
level $\gamma$ used to define the sufficiently sceptical prior:
\begin{align}
  \label{eq:success}
  \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)
  \leq \BFcol{0}{\mathrm{S}}(\hat{\theta}_o; \gy) =\gamma,
\end{align}
we have established \emph{replication success at level $\gamma$}.

For example, if we observe $z_o = 3$ (equivalent to minimum Bayes factor
$\minBF_{o} = 1/18$) and choose a level $\gamma = 1/10$
the sufficiently sceptical relative prior variance \eqref{ggamma} is
$\gy = 1.6$. If a replication is conducted with the
same precision ($c = 1$) and we observe $z_r = 2.5$ (equivalent to
minimum Bayes factor $\minBF_{r} = 1/5.5$ and relative effect
estimate $d = z_{r}/(z_{o}\sqrt{c}) = 0.83$), using equation
\eqref{eq:BFr} this would lead to $\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r;
1.6) = 1/3.5$, which means
that the replication was not successful at level $\gamma = 1/10$.
However, if we had chosen a less stringent level, \eg$\gamma =
1/3$, the replication would have been considered successful
since then $\gy = 0.4$ and
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; 0.4)
= 1/7.4$.

\subsection{The sceptical Bayes factor}
Apart from specifying a level $\gamma$, the described procedure offers an
automated way to assess replication success. One way to remove this dependence
is to find the smallest level $\gamma$ where replication success can be
established. We thus call this level the \emph{the sceptical Bayes factor}
\begin{align}
  \label{eq:bfs}
  \BFs
  =
  %% \inf_{\gy \in [0, \gminBFo]}
  \inf
  \Big\{\gamma :
  \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)
  \leq \gamma\Big\},
\end{align}
and replication success at level $\gamma$ is equivalent with $\BFs \leq \gamma$.

Figure \ref{fig:bfs} shows $\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)$ and
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; \gy)$ as a function of $\gy$ for several
values of $z_o$ and $d$ along with the corresponding $\BFs$. Typically, $\BFs$
is given by the height of the intersection between
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)$ and
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; \gy)$ in $\gy$. It may also happen that
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)$ remains below
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; \gy)$ for all values of $\gy$, in such
situations $\BFs$ is equal to the original minimum Bayes factor $\minBF_o$.
Finally, in some pathological cases it may happen that either $z_o$, $d$, or
both are so small that replication success cannot be established for any level
$\gamma$ and hence $\BFs$ does not exist. This means that the replication study
was unsuccessful since it is impossible for the advocate to convince the sceptic
at any level of evidence.

\begin{figure}[!htb]
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{images/paper1/plot-BFs-1}
\end{knitrout}
\caption{Bayes factors $\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g)$ and
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ as a function of the sufficiently
sceptical relative prior variance $\gy$. In all examples
$c = \sigma^2_o/\sigma^2_r = 1$. Minimum Bayes factors $\minBF_o$ are indicated
by dots, sceptical Bayes factors $\BFs$ are indicated by crosses where existent.}
\label{fig:bfs}
\end{figure}

In terms of computing the sceptical Bayes factor, it is worth noting that for
the special case when the replication is conducted with the same precision as
the original study ($c = 1$) and $\BFs$ is located at the intersection of
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g)$ and
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ in $g$, there is an explicit expression
for $\BFs$
\begin{align}
  \label{eq:BFsc1}
  \BFs =
  \sqrt{-\frac{z^2_{o}}{k} \cdot \frac{1 + d^{2}}{2}} \cdot
  \exp\left\{-\left(\frac{z_o^2}{2} + \frac{k}{1 + d^2}\right)\right\}
\end{align}
with
$$k = \lw{}\left(-\frac{z_{o}^{2}}{\sqrt{2}} \cdot \frac{d^{2} + 1}{2} \cdot \exp\left\{-\frac{z_o^2}{2}
  \left[1 + \frac{(1 -
d)^2}{2}\right]\right\}\right),$$ see Appendix \ref{appendix:bfs} for details.

\section{Properties}
\label{sec:comparison}
To study properties of the sceptical Bayes factor and facilitate comparison with
other methods we will look at the requirements for replication success based on
the relative effect estimate $d = \that_{r}/\that_{o}$, the variance ratio
$c = \sigma^{2}_{o}/\sigma^{2}_{r}$ and the original minimum Bayes factor
$\minBF_{o}$ (respectively the original $z$-value $z_{o}$). This perspective is
helpful because it disentangles how the method reacts to changes in
compatibility of the effect estimates ($d$), evidence from the original study
($\minBF_{o}$), and the change in sample size of the replication compared to the
original study ($c$).

The condition for replication success at level $\gamma$ from \eqref{eq:success}
is equivalent to
\begin{align}
  \label{eq:RSCond-d}
  \log \left\{ \frac{1/c + 1}{(1/c + \gy)(1 + \gy)}\right\} +
  \frac{z_o^2}{1 + 1/\gy} &\leq z_o^2 \left(\frac{d^2}{1/c +
    \gy} - \frac{\left(d - 1\right)^2}{1/c + 1} \right).
\end{align}
On the right-hand side of~\eqref{eq:RSCond-d} the $Q$-statistic
\begin{align}
  \label{eq:Qstat}
  Q
  = \frac{(\hat{\theta}_o - \hat{\theta}_r)^2}{\sigma^2_o + \sigma^2_r}
  %% = \frac{(z_r - z_o\sqrt{c})^2}{1 + c}
  = \frac{z_o^2(d - 1)^2}{1/c + 1}
\end{align}
appears. The $Q$-statistic was proposed as a measure of incompatibility among
original and replication effect estimates since its distribution is known for
standard meta-analytic models of effect sizes \citep{Hedges2019}. The connection
to the sceptical Bayes factor is such that $Q$ acts as a penalty term
in~\eqref{eq:RSCond-d} and a larger value will lower the degree of replication
success possible. However, as we will see, the sceptical Bayes factor goes
beyond assessing effect estimate compatibility as there is also a trade-off with
the amount of evidence that the replication study provides against the null.

>>>>

This paper is structured as follows: Section \ref{sec1:methods} presents the
derivation of the sceptical Bayes factor. Its asymptotic and finite sample
properties are then compared with other measures of replication success in
Section \ref{sec1:comparison}. An extension to non-normal models is presented in
Section \ref{sec1:tdist}. Section \ref{sec1:applications} illustrates how the
method works in practice using case studies from the \emph{Social Sciences
  Replication Project} \citep{Camerer2018}. Section \ref{sec1:discussion}
provides concluding remarks about strengths, limitations and extensions of the
method.

@@

We adopt a meta-analytic framework and consider the effect estimates as the
data, rather than their underlying samples, and assume that
$\that_k \given \theta \sim \Nor(\theta, \sigma^2_k)$ for $k \in \{o, r\},$ \ie
normality of the effect estimates around $\theta$, with known variances equal to
their squared standard errors. For studies with reasonable sample size, this
framework usually provides a good approximation for a wide range of (suitably
transformed) effect size types \citep[Chapter 2.4]{Spiegelhalter2004}. For
example, means and mean differences (no transformation), odds ratios, hazard
ratios, risk ratios (logarithmic transformation), or correlation coefficients
(``Fisher-$z$'' transformation). We refer to the literature of meta-analysis for
details about transformations of effect sizes \citep[\eg][Chapter
11.6]{Cooper2019}. The normal model in combination with conjugate priors enables
derivation of closed-form expressions in many cases, which allows us to easily
study limiting behaviour and facilitates interpretability. In Section
\ref{sec1:tdist}, we will present relaxations of the normality assumption, which
can lead to more accurate inferences when studies have small sample sizes and/or
show extreme results.

\section{Reverse-Bayes assessment of replication success with Bayes factors}
\label{sec1:methods}
% ``judgements are made concerning the values that the final
% probabilities would have after some imaginary results are obtained; then Bayes'
% theorem is used backwards to estimate what the consistent initial probabilities
% must be assumed to be; and lastly Bayes's theorem is used (forwards) to obtain
% estimates of the final probabilities in the light of the actual observations.
% This device takes \emph{seriously} the view that a theory of probability or of
% rationality is no more than a theory of consistency'' (p. 126).
The idea of reversing Bayes' theorem was first proposed by \citet{Good1950}. He
acknowledged that in many situations there is no obvious choice for the prior
distributions involved in Bayesian analyses. On the other hand, we are often
more certain which posterior inferences would convince us regarding the
credibility of a hypothesis. For this reason, Good inverted Bayes' theorem and
derived priors, which combined with the observed data, would lead to posterior
inferences that were specified beforehand (\eg the data favour one hypothesis
over another). His reverse-Bayes inference then centred around the question
whether the resulting prior is plausible, and if so, this would legitimise the
posterior inference. See Figure~\ref{fig1:RBschmema} for a graphical illustration
of this process.

\begin{figure}[!htb]
  \centering
%% first attempt
\begin{tikzpicture}[node distance = 5em, thick]
  % nodes %
  \node[font = \large] (data) {$f(\text{data} \,|\, \theta)$};
  \node[font = \large] (prior) [left=7em of data] {$f(\theta)$};
  \node[font = \large] (post) [right=7em of data] {$f(\theta \,|\, \text{data})$};
  \node[text width=10em,text centered] (RBinference) [below of=prior] {Reverse-Bayes inference};
  \node[text width=10em, text centered] (FBinference) [below of=post] {Forward-Bayes inference};

  % edges %
  \draw [->, dotted] (prior) to [out=45, in=135, looseness=0.3] node[above]{Bayesian updating} (post);
  %% other names could be:
  %% - reverse Bayesian updating vs forward Bayesian updating
  %% - Bayesian reverting
  \draw [->, dashed] (post) to [out=225, in=315, looseness=0.3] node[below]{Bayesian downdating} (prior);
  \draw [->, dashed] (prior) to [out=270, in=90, looseness=0] (RBinference);
  \draw [->, dotted] (post) to [out=270, in=90, looseness=0] (FBinference);
\end{tikzpicture}
\caption{Schematic illustration of reverse-Bayes and forward-Bayes inference
  for an unknown parameter $\theta$.}
\label{fig1:RBschmema}
\end{figure}

@@

\subsection{Data from the original study}
For the effect estimate $\that_o \given \theta \sim \Nor(\theta, \sigma^2_o)$
from the original study consider a hypothesis test comparing the null hypothesis
$H_0\colon$ $\theta = 0$ to the alternative $\HS \colon$ $\theta \neq 0$.
Specification of a prior distribution for $\theta$ under $\HS$ is now required
for Bayesian hypothesis testing. A typical choice \citep{Jeffreys1961} is a
local alternative, a unimodal symmetric prior distribution centred around the
null value. We consider the sceptical prior
$\theta \given \HS \sim \Nor(0, \sigma^2_s = g \cdot \sigma^2_o)$ with relative
sceptical prior variance $g$ for this purpose (relative to the variance from the
original estimate $g = \sigma^2_s/\sigma^2_o$), resembling the $g$-prior known
from the regression literature \citep{Zellner1986, Liang2008}. The explicit form
of the Bayes factor is then given by
\begin{align}
  \label{eq1:BFo}
  \BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)
  = \sqrt{1 + g} \cdot
  \exp\left\{-\frac{1}{2} \cdot \frac{g}{1 + g} \cdot z^2_o\right\}.
\end{align}

The Bayes factor from equation ~\eqref{eq1:BFo} is shown in Figure ~\ref{fig1:bfo}
as a function of $g$ and for different original $z$-values $z_o$. For fixed
$z_o$, it is well known that this Bayes factor is bounded from below by
\begin{align}
  \label{eq1:minBFo}
  \minBF_{o} =
  \begin{cases}
    |z_o| \cdot  \exp(-z_o^2/2) \cdot \sqrt{e} & \text{for} ~ |z_o| > 1 \\
    1 & \text{for} ~ |z_o| \leq 1
  \end{cases}
\end{align}
which is reached at $\gminBFo = \max\{0, z^2_o - 1\}$ \citep{Edwards1963}.
Further increasing the relative sceptical prior variance increases
\eqref{eq1:BFo} indefinitely because of the Jeffreys-Lindley paradox, \ie
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g) \to \infty$ for $g \to \infty$
\citep[][Section 6.1.4]{Bernardo2000}. Hence, for a relative sceptical prior
variance $g \in [0, \gminBFo]$, the resulting Bayes factor will be
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g) \in [\minBF_o, 1]$.

\begin{figure}[!htb]
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{images/paper1/plot-BFo-1.pdf}
\end{knitrout}
\caption{Bayes factor $\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ as a function of
relative sceptical prior variance $g$ for different values of
$|z_o| = |\that_o|/\sigma_o$. Minimum Bayes factors $\minBF_{o}$ are indicated
by dots. Dashed vertical lines indicate sufficiently sceptical relative prior
variance $g_{\scriptscriptstyle \gamma}$ at level $\gamma = 1/10$, if they
exist.}
\label{fig1:bfo}
\end{figure}

@@

From Figure \ref{fig1:bfo} we can see that the more compelling the original data
(\ie the larger $|z_o|$), the smaller the sufficiently sceptical relative prior
variance $\gy$ needs to be in order to make the result no longer convincing at
level $\gamma$. In the most extreme case, when $|z_{o}| \to \infty$ and $\gamma$
remains fixed, the sufficiently sceptical prior variance will converge to
zero (Appendix \ref{appendix:lambertW}). On the other hand, if $|z_o|$ is not
sufficiently large, $\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ will either be
always increasing in $g$ (if $|z_o| \leq 1$) or it will reach a minimum above
the chosen level $\gamma$. In both cases the sufficiently sceptical relative
prior variance $\gy$ is not defined since there is no need to challenge an
already unconvincing result.

@@

\subsection{Data from the replication study}
\label{sec1:bfr}
In order to assess whether the original finding can be replicated in an
independent study, a replication study is conducted, leading to a new effect
estimate $\hat{\theta}_r$. In light of the new data, the sceptic is now
challenged by an advocate of the original finding. This is formalised with
another Bayes factor, which compares the plausibility of the replication effect
estimate $\hat{\theta}_r \given \theta \sim \Nor(\theta, \sigma^2_r)$ under the
sceptical prior $\HS \colon$ $\theta \sim \Nor(0, g \cdot \sigma^2_o)$ relative
to the advocacy prior $\HA \colon$ $\theta \sim \Nor(\that_o, \sigma^2_o)$. The
the view of an advocate is represented by $\HA$ since this is the posterior of
$\theta$ given the original estimate and a uniform prior (also the reference
prior for this model). The Bayes factor is given by
\begin{align}
  \label{eq1:BFr}
  \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g)
%%   &= \sqrt{\frac{1/c + 1}{1/c + g}} \cdot
%%   \exp\left\{-\frac{1}{2}\left(\frac{z_r^2}{1/c + g} -
%%   \frac{\left(z_r - z_o \sqrt{c}\right)^2}{1 + c} \right)\right\}
  &= \sqrt{\frac{1/c + 1}{1/c + g}} \cdot
  \exp\left\{-\frac{z_o^2}{2}\left(\frac{d^2}{1/c + g} -
  \frac{\left(d - 1\right)^2}{1/c + 1} \right)\right\}
\end{align}
so it depends on the original $z$-statistic $z_o$, the relative sceptical prior
variance $g$, the relative effect estimate $d = \that_r/\that_o$, and the
relative variance $c = \sigma^2_o/\sigma^2_r$.

Our goal is now to define a condition for \emph{replication success} in terms of
\eqref{eq1:BFr}. It is natural to consider a replication successful if the
replication data favour the advocate over the sceptic to a higher degree than
the sceptic's initial objection to the original study. More formally, we say
that if the Bayes factor from \eqref{eq1:BFr} evaluated at the sufficiently
sceptical relative prior variance $\gy$ is not larger than the corresponding
level $\gamma$ used to define the sufficiently sceptical prior:
\begin{align}
  \label{eq1:success}
  \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)
  \leq \BFcol{0}{\mathrm{S}}(\hat{\theta}_o; \gy) =\gamma,
\end{align}
we have established \emph{replication success at level $\gamma$}.

For example, if we observe $z_o = 3$ (equivalent to minimum Bayes factor
$\minBF_{o} = 1/18$) and choose a level $\gamma = 1/10$
the sufficiently sceptical relative prior variance \eqref{ggamma} is
$\gy = 1.6$. If a replication is conducted with the
same precision ($c = 1$) and we observe $z_r = 2.5$ (equivalent to
minimum Bayes factor $\minBF_{r} = 1/5.5$ and relative effect
estimate $d = z_{r}/(z_{o}\sqrt{c}) = 0.83$), using equation
\eqref{eq1:BFr} this would lead to $\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r;
1.6) = 1/3.5$, which means
that the replication was not successful at level $\gamma = 1/10$.
However, if we had chosen a less stringent level, \eg$\gamma =
1/3$, the replication would have been considered successful
since then $\gy = 0.4$ and
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; 0.4)
= 1/7.4$.

\subsection{The sceptical Bayes factor}
Apart from specifying a level $\gamma$, the described procedure offers an
automated way to assess replication success. One way to remove this dependence
is to find the smallest level $\gamma$ where replication success can be
established. We thus call this level the \emph{the sceptical Bayes factor}
\begin{align}
  \label{eq1:bfs}
  \BFs
  =
  %% \inf_{\gy \in [0, \gminBFo]}
  \inf
  \Big\{\gamma :
  \BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)
  \leq \gamma\Big\},
\end{align}
and replication success at level $\gamma$ is equivalent with $\BFs \leq \gamma$.

Figure \ref{fig1:bfs} shows $\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)$ and
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; \gy)$ as a function of $\gy$ for several
values of $z_o$ and $d$ along with the corresponding $\BFs$. Typically, $\BFs$
is given by the height of the intersection between
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)$ and
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; \gy)$ in $\gy$. It may also happen that
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; \gy)$ remains below
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; \gy)$ for all values of $\gy$, in such
situations $\BFs$ is equal to the original minimum Bayes factor $\minBF_o$.
Finally, in some pathological cases it may happen that either $z_o$, $d$, or
both are so small that replication success cannot be established for any level
$\gamma$ and hence $\BFs$ does not exist. This means that the replication study
was unsuccessful since it is impossible for the advocate to convince the sceptic
at any level of evidence.

\begin{figure}[!htb]
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{images/paper1/plot-BFs-1}
\end{knitrout}
\caption{Bayes factors $\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g)$ and
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ as a function of the sufficiently
sceptical relative prior variance $\gy$. In all examples
$c = \sigma^2_o/\sigma^2_r = 1$. Minimum Bayes factors $\minBF_o$ are indicated
by dots, sceptical Bayes factors $\BFs$ are indicated by crosses where existent.}
\label{fig1:bfs}
\end{figure}

In terms of computing the sceptical Bayes factor, it is worth noting that for
the special case when the replication is conducted with the same precision as
the original study ($c = 1$) and $\BFs$ is located at the intersection of
$\BFcol{\mathrm{S}}{\mathrm{A}}(\hat{\theta}_r; g)$ and
$\BFcol{0}{\mathrm{S}}(\hat{\theta}_o; g)$ in $g$, there is an explicit expression
for $\BFs$
\begin{align}
  \label{eq1:BFsc1}
  \BFs =
  \sqrt{-\frac{z^2_{o}}{k} \cdot \frac{1 + d^{2}}{2}} \cdot
  \exp\left\{-\left(\frac{z_o^2}{2} + \frac{k}{1 + d^2}\right)\right\}
\end{align}
with
$$k = \lw{}\left(-\frac{z_{o}^{2}}{\sqrt{2}} \cdot \frac{d^{2} + 1}{2} \cdot \exp\left\{-\frac{z_o^2}{2}
  \left[1 + \frac{(1 -
d)^2}{2}\right]\right\}\right),$$ see Appendix \ref{appendix:bfs} for details.

\section{Properties}
\label{sec1:comparison}
To study properties of the sceptical Bayes factor and facilitate comparison with
other methods we will look at the requirements for replication success based on
the relative effect estimate $d = \that_{r}/\that_{o}$, the variance ratio
$c = \sigma^{2}_{o}/\sigma^{2}_{r}$ and the original minimum Bayes factor
$\minBF_{o}$ (respectively the original $z$-value $z_{o}$). This perspective is
helpful because it disentangles how the method reacts to changes in
compatibility of the effect estimates ($d$), evidence from the original study
($\minBF_{o}$), and the change in sample size of the replication compared to the
original study ($c$).

The condition for replication success at level $\gamma$ from \eqref{eq1:success}
is equivalent to
\begin{align}
  \label{eq1:RSCond-d}
  \log \left\{ \frac{1/c + 1}{(1/c + \gy)(1 + \gy)}\right\} +
  \frac{z_o^2}{1 + 1/\gy} &\leq z_o^2 \left(\frac{d^2}{1/c +
    \gy} - \frac{\left(d - 1\right)^2}{1/c + 1} \right).
\end{align}
On the right-hand side of~\eqref{eq1:RSCond-d} the $Q$-statistic
\begin{align}
  \label{eq1:Qstat}
  Q
  = \frac{(\hat{\theta}_o - \hat{\theta}_r)^2}{\sigma^2_o + \sigma^2_r}
  %% = \frac{(z_r - z_o\sqrt{c})^2}{1 + c}
  = \frac{z_o^2(d - 1)^2}{1/c + 1}
\end{align}
appears. The $Q$-statistic was proposed as a measure of incompatibility among
original and replication effect estimates since its distribution is known for
standard meta-analytic models of effect sizes \citep{Hedges2019}. The connection
to the sceptical Bayes factor is such that $Q$ acts as a penalty term
in~\eqref{eq1:RSCond-d} and a larger value will lower the degree of replication
success possible. However, as we will see, the sceptical Bayes factor goes
beyond assessing effect estimate compatibility as there is also a trade-off with
the amount of evidence that the replication study provides against the null.

Applying some algebraic manipulations to \eqref{eq1:RSCond-d}, one can show that
replication success at level $\gamma$ is achieved if and only if the relative
effect estimate $d$ falls within a success region given by
\begin{align}
  \label{eq1:dsuccess}
  \begin{cases}
    d \not\in [M - \sqrt{A/B}, M + \sqrt{A/B}] & \text{if} ~ \gy < 1 \\
    d \geq [1 + (1 + 1/c)\{1/2 - \log(2)/z_o^2\}]/2
    & \text{if} ~ \gy = 1 \\
   d \in [M - \sqrt{A/B}, M + \sqrt{A/B}] & \text{if} ~ \gy > 1\\
  \end{cases}
\end{align}
where
\begin{align*}
  M &= \frac{1/c + \gy}{\gy - 1} \\
  A &= \log \left\{ \frac{1/c + 1}{(1/c + \gy)(1 + \gy)}\right\}\Bigm/ z_o^2 +
  \frac{\gy}{1 + \gy} + \frac{1}{1 - \gy} \\
  B &= \frac{1 - \gy}{(1/c + \gy)(1/c + 1)}.
\end{align*}
